### 2021-04-23 input/figproof.tex
wip

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:proofA}Stale $\alpha$ messages may
      stop other $\alpha$ messages from reaching all \processes ($b$
      and $c$) that require it along the shortest path. Consistent
      partitioning requires eventual purging of stale messages (see
      Figure~\ref{fig:proofB}).]  {\input{input/figproofA.tex}}
    \hspace{10pt}
    \subfloat[Part B][\label{fig:proofB}Stale $\alpha$ messages may
      stop $\delta$ messages from reaching processes with the corresponding
      partition. Consistent partitioning requires that all \processes propagate
      $\delta$ messages as often as possible (I'); detect possible inconsistencies
      when parents' $\alpha$ messages contradict history or state (III); notify
      processes of possible inconsistencies (IV).]
    {\input{input/figproofB.tex}}
    \caption{\label{fig:proof}Concurrent operations,
      delivery orders, and stale control information may impair
      consistent partitioning. Dynamic consistent partitioning
      requires eventual purging of stale messages, and best eventual
      forwarding.}
  \end{center}
\end{figure*}

>>>>

\begin{figure*}
  \begin{center}
    \subfloat[Part A][\label{fig:proofA}Stale $\alpha$ messages may
      stop other $\alpha$ messages from reaching all \processes ($b$
      and $c$) that require it along the shortest path. Consistent
      partitioning requires eventual purging of stale messages (see
      Figure~\ref{fig:proofB}).]  {\input{input/figproofA.tex}}
    \hspace{10pt}
    \subfloat[Part B][\label{fig:proofB}Stale $\alpha$ messages may
      stop $\delta$ messages from reaching processes with the corresponding
      partition. Consistent partitioning requires that all \processes propagate
      $\delta$ messages as often as possible (I'); detect possible inconsistencies
      when parents' $\alpha$ messages contradict history or state (III); notify
      processes of possible inconsistencies (IV).]
    {\input{input/figproofB.tex}}
    \caption{\label{fig:proof}Concurrent operations, delivery orders,
      and stale control information may impair consistent
      partitioning. Dynamic consistent partitioning requires eventual
      purging of stale messages, and best eventual forwarding.
      $before$, $last$, and $expect$ respectively refer to the
      delivery history, the last delivered $\alpha$ message, and
      the message expected to reach consistent partitioning.}
  \end{center}
\end{figure*}

### 2021-04-23 input/figproofB.tex
mostly about implem
wip

  \newcommand\X{85pt}
  \newcommand\Y{-50pt}

@@

  \draw[<->] (5pt - \X, 0.4*\Y) --
  node[below, font=\small]{\textbf{I': propagation}}
  (-5pt, 0.4*\Y); %% I as well

@@

  \draw[<->] (5pt , 0.4*\Y) --
  node[below, font=\small]{\textbf{III: detection}}
  (-5pt + \X, 0.4*\Y); %% IV

@@

  \draw[<->] (5pt +\X , 0.4*\Y) --
  node[below, font=\small]{\textbf{IV: propagation}}
  (-5pt + 2*\X, 0.4*\Y); %% V

>>>>

  \newcommand\X{85pt}
  \newcommand\Y{-60pt}

@@

  \draw[<->] (5pt - \X, 0.4*\Y) --
  node[below, font=\small, align=center]{\textbf{I': propagation}\\\textbf{of $\delta$ messages}}
  (-5pt, 0.4*\Y); %% I as well

@@

  \draw[<->] (5pt , 0.4*\Y) --
  node[below, font=\small, align=center]{\textbf{III: detection}}
  (-5pt + \X, 0.4*\Y); %% IV

@@

  \draw[<->] (5pt +\X , 0.4*\Y) --
  node[below, font=\small, align=center]{\textbf{IV: propagation}\\\textbf{of $\delta?$ messages}}
  (-5pt + 2*\X, 0.4*\Y); %% V

### 2021-04-23 input/proposal.tex
wip

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \process becoming a source must notify every \process that
is closer to it than any other source. Since epidemic dissemination
and scoped broadcast have well-aligned preoccupations, we assume
implementations relying on forwarding of messages from neighbor to
neighbor.

@@

\begin{proof}
  Figure~\ref{fig:proofB} depicts all possible cases that need
  appropriate purging. A chain of \processes $[d, e, f, g]$ delivered
  and forwarded a message $\alpha_d$. Some \processes may have
  delivered messages about other partitions before or after
  $\alpha_d$.
  \begin{asparadesc}
  \item [\processes with last $\alpha_d$:] \Processes such as
    \Process~$d$ where the last and best partition is still $P_d$. By
    propagating from \process to \process the corresponding $\delta_d$
    message, these \processes purge $\alpha_d$.
  \item [\processes with $\alpha_d^y \rightarrow \alpha_h^z$ with last
    $\alpha_h^z$, for $\alpha_h^z < \alpha_d^y$:] \Processes such as
    \Process~$e$ do not directly suffer from the non-delivery of
    $\delta_d$. They need to purge their own $P_h$ if need
    be. Nonetheless, they deliver and forward $\alpha_h$ that at least
    one following \process will receive:
  \item [\processes with $\delta_h \rightarrow \alpha_d$:] \Processes
    such as \Process~$f$ do not belong to the preceding category, for
    they already delivered $\delta_h$. However, best eventual
    forwarding guarantees that they receive $\alpha_h$ that
    contradicts their history or state. In such case, the detecting
    \processes purge their own $\alpha_d$. The detecting \processes
    must also notify subsequent \processes that their current
    partition \emph{may be} deleted, for their parent may have blocked
    the corresponding $\delta$ messages. We note such notification
    $\delta_d?$ to emphasize the uncertainty of a $\delta$ message.
  \item [\processes with last $\alpha_d$ receiving $\delta_d?$ from
    their parent:] \Processes such as \Process~$g$ have $\alpha_d$ and
    preceding delivered messages have not importance. Such \processes
    must trust detecting \processes by delivering and forwarding
    $\delta_d?$. Messages are subject to all aforementioned blocking
    conditions, but are also solved by aforementioned mechanisms.
  \end{asparadesc}

@@

\vfill

\TODO{spacing.}

\newpage

@@

To implement last win order as stated in Definition~\ref{def:lwo},
each \process maintains a vector of versions that associates the
respective known local counter of each known source, or has-been
source.  Each \process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  The size
of such vector increases linearly with the number of sources that the
\process ever known which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case. Unfortunately, \processes cannot
remove any entry from this vector without running a costly distributed
consensus protocol to ensure that no \process can mistake an old but
removed entry for a new one. However, following the principles of
scoped broadcast, we expect that every \process knows only a small
subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such
$\delta$ message implements the uncertain $\delta?$ message of
Figure~\ref{fig:proofB}. It fulfils the double role of removing stale
control information without modifying the local counter of the source
(since the partition may still exist), and removing loops to ensure
termination (since \processes do not modify their local state to
reflect such delivery, messages piggyback monotonically such
increasing control information). Thus, $\delta$ messages used in
Propagation~I' are more efficient both in propagation time and
traffic, for they propagate using every communication link available
in the partition, and they do not need to carry their path.
Fortunately, Propagation~I' happens more often than Propagation~IV.
Yet, \processes that belong to a partition need to save the path taken
by the corresponding $\alpha$ messages to enable Propagation~IV.  In
the worst case, the path includes every \processes in the system but
converges in average to the average diameter of partitions. Thus,
message paths and \NAME have an interesting synergy: Paths tend to be
smaller as the number of sources in the system increases. Bloom
filters could introduce another trade-off by further decreasing the
size of path data structures at the cost of premature stopped
propagation that could impair consistent
partitioning~\cite{whitaker2002forwarding}.

>>>>

Unfortunately, \processes do not share a common global knowledge of
the network state. For \processes to reach consistent partitioning
together, a \process becoming a source must notify every \process that
is closer to it than any other source. Since epidemic dissemination
and scoped broadcast have well-aligned preoccupations, we assume
implementations relying on forwarding of messages from
neighbor-to-neighbor.

@@

\begin{proof}
  Figure~\ref{fig:proofB} depicts all possible cases that need
  appropriate purging. A chain of \processes $[d, e, f, g]$ delivered
  and forwarded a message $\alpha_d$. Some \processes may have
  delivered messages about other partitions before or after
  $\alpha_d$.
  \begin{asparadesc}
  \item [\processes with last $\alpha_d$:] \Processes such as
    \Process~$d$ where the last and best partition is still $P_d$. By
    propagating from \process to \process the corresponding $\delta_d$
    message, these \processes purge $\alpha_d$.
  \item [\processes with $\alpha_d^y \rightarrow \alpha_h^z$ with last
    $\alpha_h^z$, for $\alpha_h^z < \alpha_d^y$:] \Processes such as
    \Process~$e$ do not directly suffer from the non-delivery of
    $\delta_d$. They need to purge their own $P_h$ if need
    be. Nonetheless, they deliver and forward $\alpha_h$ that at least
    one following \process will receive:
  \item [\processes with $\delta_h \rightarrow \alpha_d$:] \Processes
    such as \Process~$f$ do not belong to the preceding category, for
    they already delivered $\delta_h$. However, best eventual
    forwarding guarantees that they receive $\alpha_h$ that
    contradicts their history or state. In such case, the detecting
    \processes purge their own $\alpha_d$. The detecting \processes
    must also notify subsequent \processes that their current
    partition \emph{may be} deleted. Indeed their parent may have
    blocked the corresponding $\delta$ messages and subsequent
    \processes must purge their possibly stale state, for the sake of
    consistency.  We note such notification $\delta_d?$ to emphasize
    the uncertainty of a $\delta$ message.
  \item [\processes with last $\alpha_d$ receiving $\delta_d?$ from
    their parent:] \Processes such as \Process~$g$ have $\alpha_d$ and
    preceding delivered messages have not importance. Such \processes
    must trust detecting \processes by delivering and forwarding
    $\delta_d?$. Messages are subject to all aforementioned blocking
    conditions, but are also solved by aforementioned mechanisms.
  \end{asparadesc}

@@

% \vfill

% \TODO{spacing.}

% \newpage

@@

To implement last win order as stated in Definition~\ref{def:lwo},
each \process maintains a vector of versions that associates the
respective known local counter of each known source, or has-been
source. Each message $m$ carries its source $s$ along with its version
$c$ that we subscript $m_{s, c}$.  Each \process delivers and forwards
\begin{inparaenum}[(i)]
\item only newest messages (see Line~\ref{line:ascast_version}) that
\item improve their best known partition (see
  Line~\ref{line:ascast_better}).
\end{inparaenum}
This vector enables both Propagation~I and Propagation~I'.  The size
of such vector increases linearly with the number of sources that the
\process ever known which is the number of \processes in the system
$\mathcal{O}(V)$ in the worst case.
%% Unfortunately, \processes cannot
%% remove any entry from this vector without running a costly distributed
%% consensus protocol to ensure that no \process can mistake an old but
%% removed entry for a new one.
Nevertheless, following the principles of scoped broadcast, we expect
that every \process only acknowledges a small subset of sources.

This vector of versions constitutes a summary of deliveries. \NAME
uses it to detect possible inconsistent partitioning as depicted in
Figure~\ref{fig:proofB}. A detecting \process then propagates a
$\delta$ message conveying the unique identifier $s$ and version $c$
of the partition along with a path $\pi$ comprising the identity of
\processes that forwarded the corresponding $\alpha$ message. Such
$\delta_{s, c}^{\pi}$ message implements the uncertain $\delta?$ message
of Figure~\ref{fig:proofB}. It fulfils the double role of removing
stale control information without modifying the local counter of the
source (since the partition may still exist), and removing loops to
ensure termination (since \processes do not modify their local state
to reflect such delivery, messages piggyback monotonically such
increasing control information). Thus, $\delta$ messages used in
Propagation~I' are more efficient both in propagation time and
traffic, for they propagate using every communication link available
in the partition, and they do not need to carry their path.
Fortunately, Propagation~I' happens more often than Propagation~IV.
Yet, \processes that belong to a partition need to save the path taken
by the corresponding $\alpha$ messages to enable Propagation~IV.  In
the worst case, the path includes every \processes in the system but
converges in average to the average diameter of partitions. Thus,
message paths and \NAME have an interesting synergy: Paths tend to be
smaller as the number of sources in the system increases. Bloom
filters could introduce another trade-off by further decreasing the
size of path data structures at the cost of premature stopped
propagation that could impair consistent
partitioning~\cite{whitaker2002forwarding}.
